Long term cloud enhancement events of global solar irradiance over Thessaloniki, Greece

### 2024-06-11 article/article.Rmd
Commit 2024-06-11 09:11

According to this study, in the period 1997 -- 2005, the mean AOD at $340\,\text{nm}$
$0.403$ with a trend of $-3.8\pm0.93\,\%$ per year, corresponding to a change of
$0.0153$ per year.  Using an Ångström coefficient $\alpha = 1.6$, this translates to
a change in the Ångström coefficient $\beta=0.00272$ per year (or $\beta=0.084$ in
1997 and $\beta=0.059$ in 2005).  Simulations with uvspec for the above Ångström
coefficients, with WC of
$`r round(BR[year==1997,pw_avg_mm],1)`\,\text{mm}$ and 
$`r round(BR[year==2005,pw_avg_mm],1)`\,\text{mm}$
for 1997 and 2005 respectively and for a SZA of $55^\circ$ reveal trend of
$+0.21\,\%$ per year in $\text{GHI}_\text{ref}$. 
The SZA of $55^\circ$ was chosen as representative of all days in the year in order
to get a rough estimate of the annually averaged change in clear sky irradiance.  For
the period 2005 -- 2023 we used the mean monthly values of AOD and WC from AERONET in
a similar simulation scheme to calculate the monthly mean of clear sky irradiance,
and finally the trend of $+0.14\,\%$ per year. We applied these two long term changes
(see Figure\nobreakspace{}\ref{fig:CS-change}) to the climatological
$\text{GHI}_\text{ref}$, in order to create a more realistic representation of the
clear-sky irradiance for the whole period of study.

@@

First, we evaluated the performance of the modelled $\text{GHI}_\text{ref}$ in
relation to the measured GHI for different levels of atmospheric clearness, by using
in the RTM the monthly climatological AOD and WC, less their respective standard
deviations.
<!-- $\tau_{\text{cs}} = \tau_{500\text{nm}} - 1\sigma$ and $w_{h\text{cs}} = w_h - 1\sigma$
($\text{GHI}_\text{CSm}$). -->
These values represent typical atmosphere in Thessaloniki, with lower than average load of
aerosols and humidity, which are the main factor that attenuate the GHI, excluding
clouds. 
To create the upper limit of CE identification, we used a value relative to
clear sky irradiance
($`r 100*(C4_cs_ref_ratio-1)`\,\%$)
with an additional constant offset of
$`r C4_GLB_diff_THRES`\,\text{W}/\text{m}^2$,
as described in Equation\nobreakspace\ref{eq:CE4}).
\begin{equation}
\text{CE} : \text{E}_\text{i} > `r C4_GLB_diff_THRES` + `r C4_cs_ref_ratio` \cdot \text{E}_\text{CSm,i} \,\,[\text{W}/\text{m}^2] \label{eq:CE4}
\end{equation}
where: $\text{E}$ the measured irradiance, $\text{E}_\text{CSm}$ the selected
modelled clear sky irradiance, and $i$ each of the one-minute observation.  This is
the criterion of our CE identification.

In the process of defining the above details, we implemented an empirical method, by
manual inspection of the CE identification results, on selected days from the whole
dataset.  We used seven sets of days, with characteristics relevant to the efficiency
of the identification threshold.  These sets were random groups of about 20 to 30
days with the following characteristics:
(a) the strongest over irradiance CE events,
(b) the largest daily total over irradiance,
(c) absence of clouds (by implementing a clear sky identification algorithm as discussed in @Natsis2023),
(d) absence of clouds and absence of EC events,
(e) with at least $60\,\%$ of the day length without clouds and presence of EC events,
(h) random selected days and
(i) some manual selected days that were included during the manual inspection process.
Where it was needed, for some of the edge cases, we also used images from a sky-cam,
to further aid the decisions of the manual inspection.

@@

The definition of the CE events with this method, has a level of subjectivity. The
actual clear sky irradiance, is not known, and can only be approximated. Although,
because our main focus is to be able to identify the OIR created by the clouds, we
expect that we were able to detect all the major events of OIR due to CE.  Where,
some CE events, with very low OIR, may be not detected, these are few and their magnitude
will not effect our results.

<!-- extreme cases criterion -->
Another aspect of the CE events, that is often reported in the relative bibliography,
are cases of extreme cloud enhancement (ECE). These are cases of CE, where the
measured intensity of the irradiance, exceeds the equivalent TSI on the top of the
atmosphere, and detected according to the Equation\nobreakspace{}\ref{eq:ECE}.
\begin{equation}
\text{ECE}: \text{GHI}_\text{i} > \cos(\theta) \times E_{i\odot} / r_{i}^2
\label{eq:ECE}
\end{equation}
where: $\theta$ the solar zenith angle, $E_{\odot}$ the solar constant, $r$ the Sun -- Earth distance, and $i$ each of the one-minute observation.

>>>>

According to this study, in the period 1997 -- 2005 the mean AOD at $340\,\text{nm}$
$0.403$ with a trend of $-3.8\pm0.93\,\%$ per year, corresponding to a change of
$0.0153$ per year.  Using an Ångström coefficient $\alpha = 1.6$, this translates to
a change in the Ångström coefficient $\beta=0.00272$ per year (or $\beta=0.084$ in
1997 and $\beta=0.059$ in 2005).  Simulations with uvspec for the above Ångström
coefficients, with WC of
$`r round(BR[year==1997,pw_avg_mm],1)`\,\text{mm}$ and 
$`r round(BR[year==2005,pw_avg_mm],1)`\,\text{mm}$
for 1997 and 2005 respectively and for a SZA of $55^\circ$ reveal trend of
$+0.21\,\%$ per year in $\text{GHI}_\text{ref}$. 
The SZA of $55^\circ$ was chosen as representative of all days in the year in order
to get a rough estimate of the annually averaged change in clear sky irradiance.  For
the period 2005 -- 2023 we used the mean monthly values of AOD and WC from AERONET in
a similar simulation scheme to calculate the monthly mean of clear sky irradiance,
and finally the trend of $+0.14\,\%$ per year. We applied these two long term changes
(see Figure\nobreakspace{}\ref{fig:CS-change}) to the climatological
$\text{GHI}_\text{ref}$, in order to create a more realistic representation of the
clear-sky irradiance for the whole period of study.

@@

First, we evaluated the performance of the modelled $\text{GHI}_\text{ref}$ in
relation to the measured GHI for different levels of atmospheric clearness, by using
in the RTM the monthly climatological AOD and WC, less their respective standard
deviations.
<!-- $\tau_{\text{cs}} = \tau_{500\text{nm}} - 1\sigma$ and $w_{h\text{cs}} = w_h - 1\sigma$
($\text{GHI}_\text{CSm}$). -->
These values represent typical atmosphere in Thessaloniki, with lower than average
load of aerosols and humidity, which are the main factor that attenuate the GHI,
excluding clouds.  With this approach the simulated $\text{GHI}_\text{ref}$ should be
generally greater than the measured GHI when aerosols are more abundant.  To
compensate for this, we increased the $\text{GHI}_\text{ref}$ by 
($`r 100*(C4_cs_ref_ratio-1)`\,\%$)
with an additional constant offset of
$`r C4_GLB_diff_THRES`\,\text{W}/\text{m}^2$,
as described in Equation\nobreakspace\ref{eq:CE4}.
\begin{equation}
\text{CE} : \text{E}_\text{i} > `r C4_GLB_diff_THRES` + `r C4_cs_ref_ratio` \cdot \text{E}_\text{CSm,i} \,\,[\text{W}/\text{m}^2] \label{eq:CE4}
\end{equation}
where: $\text{E}$ the measured irradiance, $\text{E}_\text{CSm}$ the selected
modelled clear sky irradiance, and $i$ each of the one-minute observation.  This is
the criterion of our CE identification.

These values were determined through the implementation of an empirical method with
manual inspection of the CE identification results  on selected days of the whole
dataset.  We used seven sets of days with characteristics relevant to the efficiency
of the identification threshold.  These sets were random groups of about 20 -- 30
days with the following characteristics:
(a) the strongest over irradiance CE events,
(b) the largest daily total over irradiance,
(c) absence of clouds (by implementing a clear sky identification algorithm as discussed in @Natsis2023),
(d) absence of clouds and absence of EC events,
(e) with at least $60\,\%$ of the day length without clouds and presence of EC events,
(h) randomly selected days and
(i) manually selected days.
Where needed in some of the edge cases, we also used images from a sky-camera
to further aid the decision of the manual inspection.

@@

The definition of the CE events with this method has a degree of subjectivity, since
the actual clear sky irradiance, is not known, and can only be approximated.
However, this method is capable in detecting all major CE events.  Where, some CE
events, with very low OIR may be not detected, these are few with small
over-irradiance and it is unlikely that will effect significantly our results.

<!-- extreme cases criterion -->
A sub-category  of the CE events that is often discussed in the relevant bibliography,
are cases of extreme cloud enhancement (ECE). These are cases of CE, where the
measured intensity of the irradiance, exceeds the equivalent TSI on the top of the
atmosphere, and detected according to the Equation\nobreakspace{}\ref{eq:ECE}.
\begin{equation}
\text{ECE}: \text{GHI}_\text{i} > \cos(\theta) \times E_{i\odot} / r_{i}^2
\label{eq:ECE}
\end{equation}
where: $\theta$ the solar zenith angle, $E_{\odot}$ the solar constant, $r$ the Sun -- Earth distance, and $i$ each of the one-minute observation.
